The Sub-Millimetre User Reduction Facility: an iterative map-maker for \scuba

### 2012-06-12 smurf_paper.tex
abstract and intro work

\title[SMURF: an iterative map-maker for SCUBA-2]{The Sub-Millimetre User
Reduction Facility: an iterative map-maker for SCUBA-2}

@@

\begin{abstract}
  We describe the properties of data from the Submillimetre Common
  User Bolometer Array 2 (\scuba) taken during the Shared Risk
  Observing period from 2010 February 11 to 2010 March 25, and the
  production of maps using the Sub-Millimetre User Reduction Facility
  (SMURF). Our iterative approach is similar to those used to reduce
  data from other ground-based submillimetre telescopes in recent
  years, and recovers angular scales up to the size of the array
  footprint (approximately 2\,arcmin). The method is also able to
  achieve the theoretical white-noise limit of the instrument for
  point-source studies.
\end{abstract}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron. The focal planes at each wavelength are populated with 4
rectangular subarrays, consisting of $40 \times 32$ bolometers each,
and subtend a nearly contiguous $7' \times 7'$ footprint on the sky.
Ultimately, the instrument will also be equipped with a Fourier
transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre
User Reduction Facility, SMURF, a software package for reducing the
imaging data. The reduction of FTS-2 and POL-2 data will be described
at a later date once these additional instruments have been
commissioned. The details of the instrument design, performance, and
its calibration are given in two companion papers: \citet{holland2012}
and \citet{dempsey2012}.

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1000\,\micron) have helped revolutionize
several important areas of astrophysics, including: the
characterization of early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, the early stages of planet formation.  With
$\sim10000$ total detectors, \scuba\ is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Most submillimetre imaging cameras, to date, have used bolometers to
maximize sensitivities \citep[e.g.][]{griffin2002}. Each pixel
consists of a thermal absorber to collect photons that have passed
through broad-band filters, and the signal is read-out using a
thermometer. The high-frequency broad-band noise in modern systems,
such as \scuba, is limited by photon noise from the instrument and
ambient backgrounds. The low-frequency noise, however, is dominated by
sources such as slow variations in mean backgrounds (such as thermal
variations within the cryostat, and the atmosphere for ground-based
cameras), and drift in the readout electronics. Since the
low-frequency noise is largely correlated between all, or subsets, of
the bolometers in time, it can be suppressed during map-making, since
astronomical signals have the distinct property that they are fixed in
a sky reference frame (assuming they are not time-varying). If
successful, the noise in the resulting map is said to be ``background
noise limited'', meaning that the only significant remaining noise
sources are uncorrelated spatially in the map (the theoretical limit).

There are numerous ways to attack this problem, both in terms of the
data-collection method, and data processing. Historically, both radio
and submillimetre telescope systems have used ``chopping'' to perform
rapid differences between target sky locations and some reference
(e.g., another point on the sky, internal calibrators, and different
frequencies in the case of heterodyne systems). In fact, spatial
chopping was used by SCUBA, the predecessor to \scuba, to remove
low-frequency drifts \citep{holland1999}. However, chopping (in
particular spatial chopping) has the unfortunate side-effect of
filtering out information on scales larger than the chop-throw
\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
to provide a high-degree of inherent detector stability, and rapidly
slew the telescope so that astronomical signals up to large scales are
captured in a signal band at considerably higher frequency than the
residual low-frequency drift.

In the case of chopped data, ideally the resulting data are completely
free of low-frequency noise, and therefore the data samples are
statistically uncorrelated.

often used various
``chopping'' methods to remove low-frequency drifts -- rapid
differences are made between the sky signal at a given position and an
internal reference signal; the sky signal in two different bands for
heterodyne systems (frequency chopping); and a different position in
the sky \citep[a method employed by SCUBA, the predecessor of
\scuba][]{holland1999}.

 many of them evolving
out of techniques for producing maps of the large-scale anisitropies
in the Cosmic Microwave Background (CMB)

@@

Another paper worth citing? \citet{weferling2002}

@@

@@

\item As a complete alternative to the iterative map-making procedure,
cleaning by principal component analysis (PCA) is available as an
experimental pre-processing option. As described in
Section~\ref{sec:pca}, a new set of basic vectors for the bolometer
data are found which diagonalize their covariance matrix. The
brightest components can then be subtracted from the data. A single
parameter specifies the threshold on the amplitude of the eigenvalues
to be removed, as a number of standard deviations away from the mean
value. Subarrays are cleaned independently once for the full length of
each continuousy chunk of data. Given the computational expense of
this method, and initial tests which showed little improvement over
simple high-pass filtering, this method has not yet been explored in
detail with SCUBA-2 data.

@@

For future, significantly deeper SCUBA-2 maps, in which the RMS in a
PSF-smoothed map is dominated by point-source confusion, rather than
instrumental noise, a modified matched-filter will offer improved
results. See Appendix~A in \citet{chapin2011}, which shows how to
include confusion (when known \emph{a priori}) explicitly as a noise
term in the calculation of such filters.

@@

Since our goal in this section is to measure the response of the
map-maker to extended structures, we inject a simulated signal with
power at a range of scales to a relatively empty region of the map.
It is created by drawing a realization of noise within an 800\,arcsec
on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$
which is appropriate for Galactic cirrus clouds
\citep[e.g.][]{gautier1992}. It is then filtered again with a
14.5\,arcsec Gaussian to model the effect of the SCUBA-2 optical
response. The RMS of these fluctuations are then normalized to 0.002
pW so that they are comparable to the dynamic range of M17 itself. The
minimum is subtracted so that the signal is positive. Finally, half a
period of a (1+cosine)/2 function is used to roll-off the signal to
zero between radii of 300 and 400 arcsec.

@@

\item SMURF is an iterative map-maker that can produce SCUBA-2 maps on
single high-end modern desktops in a reasonable amount of time.

>>>>

\title[SMURF: an iterative map-maker for \scuba]{The Sub-Millimetre User
Reduction Facility: an iterative map-maker for \scuba}

@@

\begin{abstract}
  We describe the properties of data from the Submillimetre Common
  User Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
  Maxwell Telescope that provides simultaneous imaging capabilities at
  450 and 850\,\micron, and the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF). With nearly 5000
  bolometers in each imaging band, and an approximately 5\,arcmin
  field-of-view, \scuba\ is presently the largest submillimetre camera
  in the world, and produces data at rates an order-of-magnitude
  faster than previous instruments. We have adopted a fast, iterative
  approach to map-making that enables data reduction on single,
  modern, high-end desktop computers with execution times that are
  shorter than the observing times.  SMURF is used in an automated
  setting both at the telescope for real-time feedback to observers,
  as well as for the production of generic science products at the
  Canadian Astronomical Data Centre. A suite of tools are also
  provided for interactive data analysis.  Three detailed case studies
  are used to: (i) explore convergence properties of the map-maker
  using simple prior constraints (Uranus -- a point source); (ii)
  demonstrate that our strategy is capable of recovering angular
  scales up to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17);
  and (iii) achieve the white-noise limit of the instrument for faint
  point-source studies (blank-field survey of the Lockman Hole).
\end{abstract}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and subtend a nearly contiguous $7'
\times 7'$ footprint on the sky.  Ultimately, the instrument will also
be equipped with a Fourier transform spectrometer
\citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{bastien2005}. This paper describes the properties of
\scuba\ data that are relevant to producing maps of the imaging data,
and the Submillimetre User Reduction Facility, SMURF, a software
package for perfoming the reduction. The analysis of FTS-2 and POL-2
data will be described at a later date once these additional
instruments have been commissioned. The details of the instrument
design, performance, and its calibration are given in two companion
papers: \citet{holland2012} and \citet{dempsey2012}.

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionize
several important areas of astrophysics, including: the
characterization of early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, the early stages of planet formation.  With
$\sim10000$ total detectors, \scuba\ is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Submillimetre imaging cameras generally use bolometers to maximize
sensitivity, in which an absorber collects photons that pass through
broad-band filters, and the signal is read out-out using a thermometer
\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
like \scuba\ are limited by photon noise from the instrument and
ambient backgrounds. The low-frequency noise, however, is dominated by
sources such as slow variations in mean backgrounds (e.g., thermal
variations within the cryostat, and the atmosphere for ground-based
cameras), and drift in the readout electronics. Such noise has a power
spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the frequency at which
it is comparable to the white noise level is called the ``$1/f$
knee''. Since the low-frequency noise is largely correlated between
all, or subsets, of the bolometers in time, it can be suppressed
during map-making, since astronomical signals have the distinct
property that they are fixed in a sky reference frame (assuming they
are not time-varying). If successful, the noise in the resulting map
is said to be ``background noise limited'', meaning that the map noise
is uncorrelated spatially, and has an amplitude that scales as the
$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
flux-density (the white noise level of a bolometer in 1\,s of
integration), and $t$ is the amount of integration time in a map
pixel.

There are numerous ways to attack the problem of map production, both
in terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds); and (ii) by offering scan
patterns that cross the sky at a wide range of position angles,
ultimately observing each portion of the map at a number of different
times. Such methods are now used by virtually all existing
ground-based submillimetre cameras, in favour of ``chopping'' methods
(where the secondary is quickly moved to modulate the signal) that
were more appropriate for older instruments that had poorer
low-frequency noise performance, and were only sensivitve to modest
angular scales
\citep[e.g.][]{glenn1998,weferling2002,laurent2005,wilson2008,kovacs2008b}.

There are three general styles of map-making that are described in the
literature for bolometer data. The best maps are produced using
\emph{maximum likelihood} techniques, in which the map is expressed as
a linear combination of the data using weights that are a function of
the time-domain noise properties of the bolometers. The first good
description of this method appears in \citet{janssen1992} for the
production of maps from the COsmic Background Explorer
(COBE). However, this method can be extremely computationally
expensive, even when small numbers of bolometers are involved,
although we note that an algorithm called ``SANEPIC'' for reducing
maps from the Balloon-borne Large-Aperture Submillimeter Telescope
(BLAST) involving hundreds of bolometers and correctly incorporating
inter-bolometer noise correlations, has been implemented with
reasonable execution speeds and memory requirements
\citep[][]{patanchon2008}.

@@

@@

% Historically, both radio
%and submillimetre telescope systems have used ``chopping'' to perform
%rapid differences between target sky locations and some reference
%(e.g., another point on the sky, internal calibrators, and different
%frequencies in the case of heterodyne systems). In fact, spatial
%chopping was used by SCUBA, the predecessor to \scuba, to remove
%low-frequency drifts \citep{holland1999}. However, chopping (in
%particular spatial chopping) has the unfortunate side-effect of
%filtering out information on scales larger than the chop-throw
%\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
%to provide a high-degree of inherent detector stability, and rapidly
%slew the telescope so that astronomical signals up to large scales are
%captured in a signal band at considerably higher frequency than the
%residual low-frequency drift.

%In the case of chopped data, ideally the resulting data are completely
%free of low-frequency noise, and therefore the data samples are
statistically uncorrelated.

%often used various
%``chopping'' methods to remove low-frequency drifts -- rapid
%differences are made between the sky signal at a given position and an
%internal reference signal; the sky signal in two different bands for
%heterodyne systems (frequency chopping); and a different position in
%the sky \citep[a method employed by SCUBA, the predecessor of
%\scuba][]{holland1999}.

@@

\item As a complete alternative to the iterative map-making procedure,
cleaning by principal component analysis (PCA) is available as an
experimental pre-processing option. As described in
Section~\ref{sec:pca}, a new set of basic vectors for the bolometer
data are found which diagonalize their covariance matrix. The
brightest components can then be subtracted from the data. A single
parameter specifies the threshold on the amplitude of the eigenvalues
to be removed, as a number of standard deviations away from the mean
value. Subarrays are cleaned independently once for the full length of
each continuousy chunk of data. Given the computational expense of
this method, and initial tests which showed little improvement over
simple high-pass filtering, this method has not yet been explored in
detail with \scuba\ data.

@@

For future, significantly deeper \scuba\ maps, in which the RMS in a
PSF-smoothed map is dominated by point-source confusion, rather than
instrumental noise, a modified matched-filter will offer improved
results. See Appendix~A in \citet{chapin2011}, which shows how to
include confusion (when known \emph{a priori}) explicitly as a noise
term in the calculation of such filters.

@@

Since our goal in this section is to measure the response of the
map-maker to extended structures, we inject a simulated signal with
power at a range of scales to a relatively empty region of the map.
It is created by drawing a realization of noise within an 800\,arcsec
on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$
which is appropriate for Galactic cirrus clouds
\citep[e.g.][]{gautier1992}. It is then filtered again with a
14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
response. The RMS of these fluctuations are then normalized to 0.002
pW so that they are comparable to the dynamic range of M17 itself. The
minimum is subtracted so that the signal is positive. Finally, half a
period of a (1+cosine)/2 function is used to roll-off the signal to
zero between radii of 300 and 400 arcsec.

@@

\item SMURF is an iterative map-maker that can produce \scuba\ maps on
single high-end modern desktops in a reasonable amount of time.
